---
title: "Kentucky Basketball"
output: html_document
date: "2022-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
```

```{r mbb-uk-setup}
library(ncaahoopR); library(tidyr)
rm(list=ls())
# source('functions.R')
```

```{r mbb-troubleshooting, echo=F, include=F}
team <- 'Kentucky'
season <- '2021-22'

dat <- get_schedule(team, season)
gid <- sched$game_id[1]
pbp <-get_pbp_game(gid)

# Who's the home team? needed to get the betting line
unique(pbp$home)
# What is the betting line?
betline <- unique(pbp$home_favored_by) 
expmarg <- ifelse(unique(pbp$home)==team , 1, -1)*betline

# Did the team of interest (TOI) cover, and how close was the result to the line?
mutdat <- dat %>% subset(game_id == gid ) %>% 
  mutate(margin = team_score - opp_score) %>% 
  mutate(expmarg = expmarg) %>% 
  mutate(relmargin = margin - expmarg )


gg_wp_chart(game_id = game_id, home_col = "black", away_col = "red")
game_excitement_index(game_id, include_spread = T)

get_boxscore(game_id)




```

# Early season success, a history

My investment in regular season Kentucky basketball has dwindled over the years. Have I gotten older? Wiser, with more responsibilities that out-weigh college sports? My priors are that Coach Calipari builds new rosters each year (read: freshman) who take time to gel and learn his offense and defensive expectations. As the game has evolved to accomodate deep threes, small-ball, and stretch 4s, Coach Cal's dribble drive offense has become more obscure. My hypothesis is that early-season success at Kentucky has changed over Cal's tenure, and I predict that as time goes on, we've lost more games in the early season or, alternatively, key games have been less competitive (e.g., against good opponents) or more competitive (against bad opponents). An alternative hypothesis is that Kentucky basketball has gotten sloppier over time as new players are less and less familiar with Cal's style of play. 

First, let's look at total wins over the first two months of the season (roughly), starting in 2001

```{r mbb-uk-load-schedules}
team <- 'Kentucky'
season.list <- paste(2001:2021, 2002:2022, sep='-')
ngames <- 15

# get_schedule(team, '2014-2015')


fun.get.nwins <- function(team, season,ngames){
  
  dat <- get_schedule(team, season)

  if(all(is.na(dat))){
    return(NA) 
  } else {
  nwins <- dat %>% 
  head(ngames) %>% 
  select('record') %>%
  unlist() %>%
  last() %>%
  strsplit(.,'-') %>%
  unlist() %>%
  head(1) %>%
  as.numeric()
  
  return(nwins)
  }
}  

winlist <- sapply(1:length(season.list), function(i) fun.get.nwins(team, season.list[i], ngames))

plot(winlist, pch=16, type='b', ylab='Wins', xlab='Season', main='Wins over the first 15 games', ylim=c(0,16), xaxt='n')
axis(side=1, at=seq(1,length(season.list), by=5), season.list[seq(1,length(season.list), by=5)])

```

Ignoring the 2020-21 season, which was an outlier as far as I'm concerned, we've been pretty steady throughout Cal's tenure in terms of number of wins, with remarkable consistency the last 7 seasons (again, disregarding the 2020-21 season). So it's not total wins *per se* that makes the last several seasons less enjoyable. What about our performance? Are we having more disappointing losses and wins? Next, I'll look at our point differential compared to the betting line. 

```{r mbb-uk-win-diff-betline}
# relative performance to betting line
team <- 'Kentucky'
# season.list <- paste(2001:2021, 2002:2022, sep='-')
season.list <- paste(2006:2021, 2007:2022, sep='-')

season <- season.list[length(season.list)]

# sched <- get_schedule(team, season)
# pbp <- get_pbp(team, season)
# pbp <- pbp[order(pbp$date),]

# There seems to be some errors here
# game_ids not matching up btwn pbp and sched, so don't try to make relational databases
# let's only keep games that are listed on the schedule for the season 
# of interest. the pbp result above includes games of the same calendar year
# but of the previous season

# tmp <- pbp %>% subset(date <= max(sched$date) & date >= min(sched$date))
# asdf <-  get_pbp_game('401258790')

fun.get.relperf <- function(sched, pbp){
  
  gids <- unique(pbp$game_id)
  
  for(i in 1:length(gids)){
    # What is the betting line?
    # gid <- sched$game_id[i]
    dat <- pbp %>% subset(game_id == gids[i])
    gamedate <- unique(dat$date)
    
    if(nrow(dat) == 0){
      # error message?
      # relmargin[i] <- NA
      next
    } else {
      betline <- unique(dat$home_favored_by)
      # based on that, what's the expected margin for the TOI?
      expmarg <- ifelse(unique(dat$home)==team , 1, -1)*betline
      
      tmp <- sched %>% subset(date == gamedate ) %>% 
      mutate(margin = team_score - opp_score) %>% 
      mutate(expmarg = expmarg) %>% 
      mutate(relmargin = margin - expmarg )
      
      if(i==1){
        mutdat <- tmp
      }
      if(i>1){
        mutdat <- rbind(mutdat,tmp)
      }
    }
    
  }

  final <- merge(sched, mutdat[,c('game_id','margin','expmarg','relmargin')], by='game_id', all=T) 
  final <- final[order(final$date),]
  
  return(final)
  
}

relperf_list <- list()

season.list <- paste(2009:2021, 2010:2022, sep='-')

for(i in 1:length(season.list)){
  sched <- get_schedule(team, season.list[i])
  pbp <-  get_pbp(team, season.list[i])
  
  if(is.null(pbp)){
    message(paste0('Warning: no play by play data available for season ', season.list[i]))
    next
  } 
  if(length(unique(pbp$date)) != length(unique(sched$date))){
    message(paste0('Warning: some games play by play data not available for season ', season.list[i]))
  }
  
  pbp <- pbp[order(pbp$date),]
  relperf_list[[i]] <- fun.get.relperf(sched,pbp)
}

sapply(relperf_list, function(x) mean(x$relmargin, na.rm=T))
sapply(relperf_list, function(x) sd(x$relmargin, na.rm=T))

hist(relperf_list[[1]]$relmargin)

tmp <- get_schedule(team, season.list[21])
tmp <- get_pbp_game('401258857')
```


<!-- Game Excitement Index -->

<!-- game_exciment_index(game_id, include_spread = T) -->

<!--     include_spread: Logical, whether to include pre-game spread in Win Probability calculations. Default = TRUE. -->

<!-- Returns GEI (Game Excitement Index) for given ESPN game_id. For more information about how these win- -->


